### 2019-12-17 latex_user_manual/appendix/GPL3_SPAMDA_license.tex

\fancyhead[LO]{License of SPAMDA}
\fancyhead[RE]{License of SPAMDA}
%\chapter{\rlap{License of SPAMDA}} \label{licenseGPLSPAMDA}
\chapter{License of SPAMDA} \label{licenseGPLSPAMDA}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{License of SPAMDA}
%\fancyhead[L]{License of SPAMDA}

\parindent=0pt

\small

SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets

Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/
	\begin{quote}
	Authors: 
		\begin{quote}
			Antonio Manuel Gómez Orellana, Juan Carlos Fernández Caballero,
			Manuel Dorado Moreno, Pedro Antonio Gutiérrez Peña and 
			César Hervás Martínez.
		\end{quote}
	\end{quote}

This program is free software: you can redistribute it and/or modify it under the
terms of the GNU General Public License as published by the Free Software Foundation,
either version 3 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY;
without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
See the GNU General Public License for more details.

A copy of the license is included in the appendix entitled "GNU GENERAL PUBLIC LICENSE" \ref{GPL_LICENSE}.
If not, see <http://www.gnu.org/licenses/>.

Additional permissions under GNU GPL version 3 section 7:

1. Redistributions of source code, with or without modification, must retain
the above full copyright notice as author attributions.

2. Redistributions in binary form and/or the use of the documentation,
with or without modification, must reproduce the above full copyright notice
as author attributions in the documentation and/or materials provided with
the distribution.

3. Modified versions of source code and/or documentation, as well as binary
distributions, must be marked in reasonable ways as different from the original version.

4. Neither name of copyright holders nor the names of its contributors may be used
to endorse or promote products derived from this software for publicity purposes
without specific prior written permission.

5. Redistribution and/or use of source code, binary format and documentation,
with or without modification, could require indemnification of licensors
and authors by anyone who conveys the material (or modified versions of it)
with contractual assumptions of liability to the recipient, for any liability
that these contractual assumptions directly impose on those licensors and authors.

SPAMDA uses some external libraries. You can see their respective notices about license,
copyright and disclaimer in the following appendices. For a more complete information about
such licenses, see the distributions provided by their authors:

-Library NetCDF Java, version 4.6.10

\hspace{0.75cm}Notice of license in the appendix entitled ``NetCDF-LICENSE'' \ref{NetCDF_LICENSE}.

-Library SLF4J, version 1.7.25

\hspace{0.75cm}Notice of license in the appendix entitled ``SLF4J-LICENSE`` \ref{SLF4J_LICENSE}.

-Library WEKA, version 3.8.1

\hspace{0.75cm}Notice of license in the appendix entitled ''WEKA-LICENSE'' \ref{WEKA_LICENSE}.

Contact information:

Juan Carlos Fernández Caballero, PhD.

email: jfcaballero@uco.es

Address: University of Cordoba, Department of Computer Science
and Numerical Analysis, AYRNA Research Group, Rabanales Campus,
Einstein Building, 3rd floor. Road Madrid-Cadiz, Km 396-A.
14071 - Cordoba (Spain).