### 2019-12-17 latex_user_manual/appendix/WEKA_license.tex

\fancyhead[LO]{WEKA-LICENSE}
\fancyhead[RE]{WEKA-LICENSE}
%\chapter*{\rlap{WEKA-LICENSE}} \label{WEKA_LICENSE}
%\phantomsection  % so hyperref creates bookmarks
%\addcontentsline{toc}{chapter}{WEKA-LICENSE}
\chapter{WEKA-LICENSE} \label{WEKA_LICENSE}

\parindent=0pt

\small

Weka (Waikato Environment for Knowledge Analysis) is a Java tool that
incorporates several standard ML techniques into a software "workbench".

WEKA version 3.8.1 - weka.jar
Available on https://www.cs.waikato.ac.nz/ml/weka/

Copyright (C) 1998-2016  University of Waikato

The core WEKA system is distributed under the GNU public license.
Please read the the section entitled ``GNU GENERAL PUBLIC LICENSE`` \ref{GPL_LICENSE}.

Packages may be distributed under various licenses - check the
description of the package in question for license details.