### 2019-12-17 latex_user_manual/chapters/acknowledgments.tex

\chapter{Acknowledgments}

	\begin{onehalfspace}

		This work has been partially subsidised by the projects TIN2017-85887-C2-1-P and TIN2017-90567-REDT of the Spanish Ministry of Economy and Competitiveness (MI\-NE\-CO), and FEDER funds of the European Union. We also thank to NVIDIA Corporation for the transfer of computational resources for research works.

		The authors thank to NOAA/OAR/ESRL PSD, Boulder, Colorado, USA for the NCEP Reanalysis data provided from their Web site at https://www.esrl.noaa.gov/psd/, and to NOAA/NDBC by its data that were collected and made freely available.

		We also thank to University of Waikato for the Weka (Waikato Environment for Knowledge Analysis) software tool, to University Corporation for Atmospheric Research/Unidata for the NetCDF (network Common Data Form) Java library and to QOS.ch for the SLF4J (Simple Logging Facade for Java) library.

	\end{onehalfspace}