### 2019-12-17 latex_user_manual/chapters/ns_installation.tex

\chapter{Installing SPADNN}

	\begin{onehalfspace}

		The process of installation of SPADNN is quite easy and it does not require administrator permissions to carry it out. Once the compressed file with SPADNN has been saved on the computer, the installation process can be performed just doing the following steps:

			\begin{enumerate}

				\item Create a folder where holding the software tool.
				\item Copy the compressed file inside the folder.
				\item Decompress the file by double click on it.

			\end{enumerate}

		After performing the above steps SPADNN has been installed on the computer. Therefore, and in order to run it, just open the ``dist'' folder and type the following sentence on the command line:

		\texttt{java -jar "SPADNN"}

		\section{How to uninstall?}

			Note that this action will remove SPADNN from your computer and any information entered or generated by the software tool will be lost.

			To uninstall SPADNN just delete the folder in which the installation process was carried out.

	\end{onehalfspace}