### 2019-12-17 latex_user_manual/chapters/ns_requirements.tex

\chapter{System requirements}

	\begin{onehalfspace}

		SPADNN has been developed in Java and it is a multi-platform software tool. In this way, any computer able to install and run a Java Virtual Machine (JVM) will be enough for running the developed tool. As SPADNN has been compiled for Java JRE 1.8v the Java version 8 needs to be installed in the system, which can be downloaded from [\textcolor{red}{poner enlace}] choosing the correct distribution depending on the system platform.

	\end{onehalfspace}