### 2019-12-17 latex_user_manual/license/licenseDoc.tex

\thispagestyle{empty}

\small

\vspace*{\fill} % El texto aparecerá en la parte inferior de la página.

\begin{quote}

	SPAMDA: Software for Pre-processing and Analysis of Meteorological DAta to build datasets.

	This is the version 1.0 of the SPAMDA manual.

	Copyright (c) 2017-2018 by AYRNA Research Group. https://www.uco.es/ayrna/

	\hspace{1.005cm}Authors:

	\hspace{1.30cm}Antonio Manuel Gómez Orellana, Juan Carlos Fernández Caballero,

	\hspace{1.30cm}Manuel Dorado Moreno, Pedro Antonio Gutiérrez Peña and

	\hspace{1.30cm}César Hervás Martínez.

	\begin{quote} % El texto aparecerá con un margen izquierdo y derecho de 1cm.
		Permission is granted to copy, distribute and/or modify this document 
		under the terms of the GNU Free Documentation License, Version 1.3 or any later
		version published by the Free Software Foundation; with no Invariant Sections, 
		with no Front-Cover Texts, with the Front-Cover logo, and with no Back-Cover Texts.
		\newline
		\newline
		A copy of the license is included in the appendix entitled "GNU Free Documentation License" \ref{GFDL_LICENSE}. If not, see <http://www.gnu.org/licenses/>.

		\vspace{0.50cm}
		Contact information:

		\begin{quote}
			Juan Carlos Fernández Caballero, PhD.

			email: jfcaballero@uco.es

			Address: University of Córdoba, Department of Computer Science and Numerical Analysis, Rabanales Campus, AYRNA Research Group, Einstein Building, 3rd floor. Road Madrid-Cádiz, Km 396-A. 14071 - Córdoba (Spain)
		\end{quote}

	\end{quote}

\end{quote}